An Answer Set Programming Framework for Reasoning about Agents' Beliefs and Truthfulness of Statements

### 2020-06-17 main.tex
Typo in (11).

\item We have adequate knowledge about actions  and their effects (e.g., the action of buying a car requires that the agent has money and its execution will result in the agent owning a car). This knowledge is represented by an action theory $Act$ in a suitable logic that allows reasoning about actions' effects and consequent changes to the world. 
        Let $\Phi_{Act}: \Sigma \times A \rightarrow 2^\Sigma$ denote a suitable transition function describing the evolution of the dynamic system of interest. $\hat\Phi_{Act}$ denotes the natural extension of $\Phi_{Act}$ to a history. %A resulting sequence $<\sigma_0,a_0,\sigma_1,a_1,\ldots,\sigma_k>$ is called a \emph{possible path}. 
        %Let us denote with $\models_A$ the entailment relation defined within the logical framework $A$ used to describe $Act$.

\item We have commonsense knowledge about how an agent of interest forms its beliefs (e.g., an agent may believe that a person attending an expensive school normally comes from a rich family, a person obtaining need-based scholarship s  usually comes from a poor family).
This knowledge is represented by a default theory with preferences $\textit{Def}$,  that enables reasoning about the state of the agent's beliefs. Let us denote with $\models_D$ the entailment relation defined over the default theory framework defining $\textit{Def}$.

@@

A set of rules $RD$ is \emph{consistent} if, for any consistent set of fluent literals $X$, the logic program 
$\{b \leftarrow body  \mid rule(r, b, body) \in RD\} \cup \{l \leftarrow \:\mid\: l\in X\}$
is consistent.
%
\begin{definition}
A $KB = \langle O, RD \rangle$ is \emph{consistent} if $O$ and $RD$ is consistent. 
\end{definition} 
%
Our goal is to identify the agent's beliefs and to use them to determine whether  a statement $stm(p, s)$ is true at a time step $t$ given $KB$. We will develop a program $\Pi(KB)$ to answer this question. We assume a finite horizon of steps, denoted by $s(0), \ldots, s(k)$. We use $h(L,S)$ to encode that fluent literal $L$ is true at step $S$. We use $believes(B,S)$ to state that the agent believes $B$ at step $S$. Furthermore, to simplify the use of $\Pi(KB)$ with current ASP solvers, we  encode a rule of the form \eqref{rule}  using {\bf (a)} the ASP atom $rule(r, head, name)$ and
{\bf (b)} the set of ASP atoms $\{mbr(l, name) \mid l \in body_{r}\}$, where \(name\) is a unique identifier for the rule; similar encodings are used for  \eqref{default} and \eqref{pref}. 
This method of encoding is similar to what is used  in answer set planning and allows for the separation between  the encoding of the problem and the rules for reasoning with it.
The main challenge to be addressed lies in the interaction between observations and defaults. 
We apply the following principles:
%
\begin{itemize}
\item[]{\bf (O)}ptimistic: The most recent observation reflects the true state of the world;
\item[]{\bf (SK)}eptical: When conflicting conclusions can be drawn then do not believe in any.
\end{itemize}

@@

The first set of rules is used to  reason about observations of the form 
\eqref{obs}:
%
\begin{eqnarray}  
h(L, S)  & \leftarrow &   obs(L,S_1), s(S),  S_1 \le S, \label{lp:obs} \\
&&\naf nr(L, S_1, S).    \nonumber\\
%               \naf \:not\_most\_recent(P, S_1, S).  \nonumber \\
% not\_most\_recent(P, S_1, S) \leftarrow s(S_1),  \label{lp:obs1}  \\ 
  nr(L, U, S)  & \leftarrow &  s(U), neg(L,L_1), obs(L_1, V), s(S), \label{lp:obs1} \\
  &&U {<} V, V {<} S.  \nonumber  %\\ 
%obs(\overline{P}, S_2), s(S),  S_1 < S_2, S_2 < S. \nonumber
\end{eqnarray}

These rules encode the principle {\bf (O)}. Rule \eqref{lp:obs1} states that $nr(L,S_1)$ is true if there exists a more recent observation of  $\neg L$. Rule \eqref{lp:obs} indicates that an observation stays true if every conflicting observation is `older'.

 For reasoning about statements of the form \eqref{rule}, we have:  %
%
\begin{equation}\label{lp:rule}
\begin{array}{ll}
believes(H, S) \leftarrow & s(S),  rule(D, H, M),  \\
        & N  == \# count \{L : mbr(L, M)\},\\
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.8in}h(L, T) \}== N_h,  \\ 
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.44in}believes(L, T) \}== N_b, \\
        & N_h+N_b==N.  \\ 
%
%& N = \# count \{L : mbr(L, M),  h(L, S) \},\\ 
%     & N  = \# count \{L : mbr(L, M)\}.
      \end{array}
\end{equation}

%
Intuitively, (\ref{lp:rule}) states that, if the body of a $rule$ statement is satisfied, then its head must be true. Recall that, according to the syntax given earlier, the head is a belief -- hence the expression $believes(H,S)$ in (\ref{lp:rule}). The `$==$' symbol denotes identity at the level of ground terms.\footnote{The variables involved are replaced by a constant during the grounding process and the identity is checked during the solving phase.} Checking whether the body is satisfied is compactly achieved by leveraging the $\#count$ aggregate of the ASP Core 2 language\footnote{Details can be found in \cite{ASPCore2Standard}}, which calculates the cardinality of a set. For instance, the expression $\# count \{L : mbr(L, M), h(L, T) \}$ finds number of literals that are in the body of the rule ($mbr(L,M)$) and hold at time step $T$ ($h(L,T)$). The combination of aggregates present in (\ref{lp:rule}) checks that the number of literals that hold or are believed equals the size of the body of the rule.\footnote{Recall that fluent symbols and belief symbols are disjoint. }

 For reasoning about defaults  of the form \eqref{default} we have:
% 
%%\iffalse
%{\small
\begin{eqnarray}
app(D, H, S) & \leftarrow  &  s(S), default(D, H, M), \nonumber \\
        & & N  == \# count \{L : mbr(L, M)\},    \nonumber\\
        & & \# count \{L : mbr(L, M),  \nonumber\\
        & & \hspace*{.8in}h(L, T) \}== N_h,  \nonumber\\ 
        & & \# count \{L : mbr(L, M), \nonumber\\
        & & N_h+N_b==N. \label{lp:app} \\ 
%     & & NC  = \# count \{L : mbr(L, M)\}, \nonumber \\ 
%    & & \# count \{L : mbr(L, M),     h(L, T) \} == NC. \nonumber \\ 
ab(D, H, S)& \leftarrow &   s(S), app(D, H, S), \nonumber\\
    & &  neg(H, H_1), app(D_1, H_1, S),  \nonumber \\
    &&   \naf defeated(D_1, H_1, S).  \label{lp:ab}      \\
defeated(D, H, S) & \leftarrow  & s(S), app(D, H, S), \nonumber\\
        & &  neg(H,H_1), \nonumber \\
        & &  believes(H_1, S).  \:\:\:  \label{lp:defeat0} \\ 
defeated(D, H, S) & \leftarrow  & s(S),  app(D, H, S),    \nonumber \\ 
        & &  neg(H,H_1), \nonumber \\ %
        & & app(D_1, H_1, S), \nonumber\\
        & & prefer(D_1, D, S).   \label{lp:defeat}\\
believes(H, S) & \leftarrow & app(D, H, S), \nonumber \\
        & & \naf ab(D, H, S), \nonumber\\
        & & \naf defeated(D, H, S).  \label{lp:holds}
%    
%defeated(D, H, S)& \leftarrow &  s(S), app(D, H, S), h(\overline{H}, S).\label{lp:defeat0} \\
%    &&  app(D, H, S),  \nonumber \\
%    &&  h(\overline{H}, S). \nonumber \\       
%defeated(D, H, S)& \leftarrow  & s(S),  app(D, H, S), \label{lp:defeat} \\
%    && app(D_1, \overline{H}, S), prefer(D_1, D, S).  \nonumber \\
%    &&  prefer(D_1, D, S).  \nonumber      \\
%h(l_1, S), \ldots, h(l_k, S), \naf ab(d, l, S)   \label{lp:def}\\
%b(l, S) \leftarrow b(l_1, S), \ldots, b(l_k, S), \naf ab(d, l, S)   \label{lp:def_bel}\\
%app(d, l, S) \leftarrow h(l_1, S), \ldots, h(l_k, S) 
%h(H, S)& \leftarrow  app(D, H, S), \naf ab(D, H, S),  \label{lp:holds} \\
% \naf ab(D, H, S), \nonumber \\ 
% & \naf defeated(D, H, S). \nonumber 
\end{eqnarray} 
%} 
%%\fi
%
\iffalse
{
        \[
        \begin{array}{llr}
     app(D, H, S)  \leftarrow &  s(S), default(D, H, M),  \nonumber\\
        & N  = \# count \{L : mbr(L, M)\},  & \label{lp:app} \\
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.8in}h(L, T) \}= N_h,  \\ 
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.44in}believes(L, T) \}= N_b, \\
        & N_h+N_b=N.  \\ 
        ab(D, H, S) \leftarrow  & s(S), app(D, H, S), neg(H,H_1), & (10)\\   %\label{lp:ab}
        &  app(D_1, H_1, S), \\
        & \naf defeated(D_1, H_1, S).        \\
        defeated(D, H, S) \leftarrow  & s(S), app(D, H, S), neg(H,H_1), \\
        & believes(H_1, S).  \:\:\: & (11) \\ %\label{lp:defeat0} 
        defeated(D, H, S) \leftarrow  & s(S),  app(D, H, S), neg(H,H_1), & (12) \\  %\label{lp:defeat}
        & app(D_1, H_1, S), \\
        & prefer(D_1, D, S).   \\
        believes(H, S) \leftarrow & app(D, H, S), \naf ab(D, H, S),\\
        & \naf defeated(D, H, S).  & (13) \\ %\label{lp:holds}
        \end{array}
        \]
        }
%
\fi

@@

 \setcounter{equation}{13}

 For reasoning about preferences of the form \eqref{pref}, we employ the rules:
% 
{
\begin{equation}\label{lp:pref}
\begin{array}{ll}
prefer(D_1, D_2, S) \leftarrow &  s(S),   prefer(D_1, D_2, M),   \\
     & N  = =\# count \{L : \\
     &  \hspace*{.3in} mbr(L, M)\},  \\ 
     & \# count \{L : mbr(L, M),  \\
     &  \hspace*{.3in}  h(L, S) \} {=} =N_h, \\
     & \# count \{L : mbr(L, M), \\
     & \hspace*{.3in} believes(L, S) \} {=} =N_b, \\
     & N_h+N_b==N.
      \end{array}
\end{equation} 
}
%
This rule defines when a preference among two defaults can be applied. 
%
%\iffalse 
%Rule~\eqref{lp:app} defines when a default is \emph{applicable}, i.e., when its body is satisfied. The interaction between defaults and rules in the knowledge base are dealt with using rules~\eqref{lp:ab}---\eqref{lp:defeat}. Rule~\eqref{lp:ab} enforces the principle {\bf (SK)}. It says $ab(d, h, s)$ that if there is another default with the conflicting conclusion ($\overline{h}$) 
%which is not defeated then $ab(d, h, s)$ is true. This atom will be used to suppress the conclusion of $d$.   
%\eqref{lp:defeat0} dismisses the applicability of a default if the contrary of its conclusion is already established. On the other hand, rule \eqref{lp:defeat} expresses that a default is defeated if there is a more preferred default with conflicting conclusion that is applicable.  
%
%
% 
%
%\item Rules for reasoning about beliefs  
%\begin{align}
%b(L, S) \leftarrow h(L, S)  \label{lp:belief}   
%\end{align} 
%
%
%\item For interacting between defaults, we have the following rules: 
%
%%\memo{
%%}
%
%\begin{center} 
%  \begin{tikzpicture} 
%%    \tikzstyle{every node}=[draw,shape=circle} 
%%    \node (s1) at (0:0) {$\alert{s_3:}$}   
%%      \draw[fill=blue!20,style=dashed] (-1.0,-1.0) rectangle (3.0,3.0)  ;
%
%      \draw (0,0) circle(0.4) node {$\textcolor{red}{\small s{-}1}$} ;
%      \draw  (2,0) circle(0.4) node  {$\textcolor{red}{\small s}$} ;
%     \draw  (4,0) circle(0.4) node {$\textcolor{red}{\small s{+}1}$} ;  
% 
%     \draw[-,style=dashed] (-1.6,0) -- (-0.4,0 )  ; 
%     \draw[-] (0.4,0) -- (1.6,0)  ;  
%     \draw[-] (2.4,0) -- (3.6,0 )  ; 
%     \draw[-,style=dashed] (4.4,0) -- (5.6,0 )  ; 
% 
%  \end{tikzpicture} 
%\end{center} 
%
%
%\begin{align}\label{lp:ab} 
%ab(D_1, L, S) \leftarrow \: \: &     default(D_1, L, Body_1),   \\
%    &      app(D_1, L, S), \nonumber \\
%    &      default(D_2, \overline{L}, Body_2), \nonumber \\
%    &      app(D_2, \overline{L}, S), \nonumber \\
%    &      \naf app(D_2, S-1).  \nonumber
%\end{align} 
%
%This rule encodes the principle {\bf (P3)}. 
%
%\fi
%
%
%
%% We will assume that things behave in accordance to the law of inertia. So, we will add to the the program the rule 
%%
%% \begin{align}
%% h(l, S+1) \leftarrow h(l, S),  \naf h(\overline{l}, S)   \label{lp:inertial} \\
%% b(l, S+1) \leftarrow b(l, S),  \naf b(\overline{l}, S)   \label{lp:inertial_bel} 
%% \end{align} 
%
%

@@

%
\begin{proposition}
For each consistent $KB$,  there exists no answer set of $\Pi(KB)$ that contains $believes(l, t)$ and $believes(\overline{l},t)$ for some 
literal $l$ and time step $t$. 
\end{proposition} 
%
\noindent
{\bf Proof.}
The proof of this proposition relies on the fact that $believes(l, t)$ is in an answer set $A$ if the logic program  rule \eqref{lp:rule} (or \eqref{lp:holds}) is satisfied by $A$. If a rule \eqref{lp:rule} is satisfied by $A$ then a $KB$ contains a rule $rule(r,l,body)$ and $body$ is satisfied by $A$. Because of the consistency of $KB$, we know that there exists no rule $rule(r', \overline{l}, body')$ such that $body'$ is satisfied by $A$. For, otherwise, $KB$ is inconsistent. This will imply that for every default $default(d, \overline{l}, body')$, if the default is applicable, then the body of the rule \eqref{lp:defeat0} is satisfied and therefore $defeated(d)$ belongs to $A$. Consequently, \eqref{lp:holds} is not satisfied for the default $d$, and hence, $believes(\overline{l}, t)$ cannot belong to $A$.

@@

Intuitively, $KB  \models {+}stm(b,s)@t$ (resp. $KB  \models {-}stm(b,s)@t$) says that at the time step $t$, at the same time step or later than the time step that the statement about $b$ is made, the statement is believed to be true (resp. believed to be false). $KB \not\models {\pm}stm(b,s)@t$ states that there is no information that supports or denies the statement at the time step $t$. Observe that $l$, $t$, and $s$ are constants representing the question whether or not $stm(b,s)$ is believed to be true at step $t$. As such, requiring $believes(b,t)$ to be in every answer set of the program 
does not mean that $believes(b,t)$ belongs to every answer set for every $t$ such that $s \le t \le k$. 
Observe that this entailment can be computed in ASP by the following rules: 
%
{
\begin{eqnarray}
\mathtt{t}(H, T) & \leftarrow & stm(H, S), s(T), T \ge S, \nonumber \\
        &  & believes(H, T). \label{q1} \\ 
\mathtt{f}(H, T) & \leftarrow & stm(H, S), s(T), T \ge S, neg(H,H_1),  \nonumber\\ 
& & believes(H_1, T).  \label{q2}\\
\mathtt{u}(H, T) & \leftarrow & stm(H, S), s(T), T \ge S, neg(H,H_1), \nonumber  \\
& & \naf believes(H, T), \nonumber \\
&&  \naf believes(H_1, T). \label{q3}
\end{eqnarray}
}
%
Let $\Pi_Q = \Pi(KB) \cup \{\eqref{q1}{-}\eqref{q3}\}$. It can be shown that 
 $KB  \models {+}stm(b,s)@t$, $KB  \models {-}stm(b,s)@t$, and  $KB  \models {\pm}stm(b,s)@t$ 
 correspond to $\Pi_Q \models \mathtt{t}(b,t)$, $\Pi_Q \models \mathtt{f}(b,t)$, and 
 $\Pi_Q \models \mathtt{u}(b,t)$, respectively. Hence, we can compute the truthfulness of a statement 
using two calls to an ASP-solver.

@@

  \begin{equation}\label{lp:rule:tt}
\begin{array}{l}
h(H, S+T) \leftarrow  s(S),  
        rule(D, H^T, M),  \\ 
     \hspace{.1cm} N = \# count \{L^T : mbr(M, L^T)\},  \\ 
    \hspace{.1cm}  \# count \{L^T : mbr(M, L^T),  h(L, T{+}S) \} {==} N. 
      \end{array}
\end{equation}

>>>>

\item We have adequate knowledge about actions  and their effects (e.g., the action of buying a car requires that the agent has money and its execution will result in the agent owning a car). This knowledge is represented by an action theory $Act$ in a suitable logic that allows reasoning about actions' effects and consequent changes to the world. 
        Let $\Phi_{Act}: \Sigma \times A \rightarrow 2^\Sigma$ denote a suitable transition function describing the evolution of the dynamic system of interest. $\hat\Phi_{Act}$ is the usual extension of $\Phi_{Act}$ to action sequences with $\hat\Phi_{Act}(\sigma, []) = \sigma$ and 
$\hat\Phi_{Act}(\sigma, [a; \beta]) = \bigcup_{\sigma' \in \Phi_{Act}(a, \sigma)} \hat\Phi_{Act}(\sigma', \beta)$ where $\sigma \in \Sigma$ and $\beta$ is an action sequence  and $a$ is executable in $\sigma$.                 
%        denotes the natural extension of $\Phi_{Act}$ to sequence. %A resulting sequence $<\sigma_0,a_0,\sigma_1,a_1,\ldots,\sigma_k>$ is called a \emph{possible path}. 
        %Let us denote with $\models_A$ the entailment relation defined within the logical framework $A$ used to describe $Act$.

\item We have commonsense knowledge about how an agent of interest forms its beliefs (e.g., an agent may believe that a person attending an expensive school normally comes from a rich family, a person obtaining need-based scholarship  usually comes from a poor family).
This knowledge is represented by a default theory with preferences $\textit{Def}$,  that enables reasoning about the state of the agent's beliefs. Let us denote with $\models_D$ the entailment relation defined over the default theory framework defining $\textit{Def}$.

@@

A set of rules $RD$ is \emph{consistent} if, for any consistent set of fluent literals $X$, the logic program 
$\{b \leftarrow body  \mid rule(r, b, body) \in RD\} \cup \{l \leftarrow \:\mid\: l\in X\}$
is consistent.
%
\begin{definition}
A $KB = \langle O, RD \rangle$ is \emph{consistent} if $O$ and $RD$ are consistent. 
\end{definition} 
%
Our goal is to identify the agent's beliefs and to use them to determine whether  a statement $stm(p, s)$ is true at a time step $t$ given $KB$. We will develop a program $\Pi(KB)$ to answer this question. We assume a finite horizon of steps, denoted by $s(0), \ldots, s(k)$. We use $h(L,S)$ to encode that fluent literal $L$ is true at step $S$. We use $believes(B,S)$ to state that the agent believes $B$ at step $S$. Furthermore, to simplify the use of $\Pi(KB)$ with current ASP solvers, we  encode a rule of the form \eqref{rule}  using {\bf (a)} the ASP atom $rule(r, head, name)$ and
{\bf (b)} the set of ASP atoms $\{mbr(l, r) \mid l \in body_{r}\}$, where \textit{name} is the name for the rule and $mbr$ stands for `\textit{member\_of}'; similar encodings are used for  \eqref{default} and \eqref{pref}. 
This method of encoding is similar to what is used  in answer set planning and allows for the separation between  the encoding of the problem and the rules for reasoning with it.
The main challenge to be addressed lies in the interaction between observations and defaults. 
We apply the following principles:
%
\begin{itemize}
\item[]{\bf (O)}ptimistic: The most recent observation reflects the true state of the world;
\item[]{\bf (SK)}eptical: When conflicting conclusions can be drawn then do not believe in any.
\end{itemize}

@@

The first set of rules is used to  reason about observations of the form 
\eqref{obs}:
%
\begin{eqnarray}  
  nr(L, U, S)  & \leftarrow &  s(U), neg(L,L_1), obs(L_1, V), s(S), \nonumber  \\
  &&U {<} V, V {<} S.   \label{lp:obs1} \\
%obs(\overline{P}, S_2), s(S),  S_1 < S_2, S_2 < S. \nonumber
h(L, S)  & \leftarrow &   obs(L,S_1), s(S),  S_1 \le S, \nonumber  \\ 
&&\naf nr(L, S_1, S).    \label{lp:obs} 
%               \naf \:not\_most\_recent(P, S_1, S).  \nonumber \\
% not\_most\_recent(P, S_1, S) \leftarrow s(S_1),  \label{lp:obs1}  \\ 
\end{eqnarray}

These rules encode the principle {\bf (O)}. Rule \eqref{lp:obs1} states that $nr(L,U,S)$ is true if there exists a more recent observation of  $\neg L$. Rule \eqref{lp:obs} indicates that an observation stays true if every conflicting observation is `older'.

 For reasoning about statements of the form \eqref{rule}, we have\footnote{
 We would like to thank the anonymous reviewer for the suggestion on how to encode this rule for better performance.  
%%believes(H,S) :- s(S), rule(D,H,M),
%%#sum{1,L : mbr(L,M);
%%-1,L : mbr(L,M), h(L,T);
%%-1,L : mbr(L,M), believes(L,T)} == 0. 
 }:  %
%
\begin{eqnarray}
h\_body(M,S) & \leftarrow  &  s(S), \nonumber \\ 
        & & N  = \# count \{L : mbr(L, M)\},    \nonumber\\
        & & N_h= \# count \{L : mbr(L, M),  \nonumber\\
        & & \hspace*{.8in}h(L, T) \},  \nonumber\\ 
        & & N_b = \# count \{L : mbr(L, M), \nonumber\\
        & &\hspace*{.44in}believes(L, T) \}, \nonumber \\
        & & N_h+N_b==N. \label{lp:body} \\
believes(H, S) & \leftarrow & s(S),  rule(R_, H, M), \nonumber \\
        && h\_body(M,S)   \label{lp:rule}
%%        & N  = \# count \{L : mbr(L, M)\},\\
%%        & N_h = \# count \{L : mbr(L, M),\\
%%        &\hspace*{.8in}h(L, T) \},  \\ 
%%        & N_b = \# count \{L : mbr(L, M),\\
%%        &\hspace*{.44in}believes(L, T) \}, \\
%%        & N_h+N_b==N.  \label{lp:believes}   
%
%& N = \# count \{L : mbr(L, M),  h(L, S) \},\\ 
%     & N  = \# count \{L : mbr(L, M)\}.
\end{eqnarray} 
%
\eqref{lp:body} defines an auxiliary predicate \textit{h\_body} where for each set $M$ of literals and time step $S$, 
$h\_body(M,S)$ is true when all literals in $M$ are true or believed at $S$. 
The `$==$' symbol denotes identity at the level of ground terms.\footnote{The variables involved are replaced by a constant during the grounding process and the identity is checked during the solving phase.} Checking whether the body is satisfied is compactly achieved by leveraging the $\#count$ aggregate of the ASP Core 2 language\footnote{Details can be found in \cite{ASPCore2Standard}}, which calculates the cardinality of a set. For instance, the expression $\# count \{L : mbr(L, M), h(L, T) \}$ finds the number of literals that are in the body of the rule ($mbr(L,M)$) and hold at time step $T$ ($h(L,T)$). The combination of aggregates present in (\ref{lp:rule}) checks that the number of literals that hold or are believed equals the size of the body of the rule.\footnote{Recall that fluent symbols and belief symbols are disjoint. }
Intuitively, (\ref{lp:rule}) states that, if the body of a $rule$ statement is satisfied, then its head must be true. Recall that, according to the syntax given earlier, the head is a belief -- hence the expression $believes(H,S)$ in (\ref{lp:rule}).

 For reasoning about defaults  of the form \eqref{default} we have:
% 
%%\iffalse
%{\small
\begin{eqnarray}
app(D, H, S) & \leftarrow  &  s(S), default(D, H, M), \nonumber \\
&& h\_body(M,S)  \label{lp:app} \\ 
%     & & NC  = \# count \{L : mbr(L, M)\}, \nonumber \\ 
%    & & \# count \{L : mbr(L, M),     h(L, T) \} == NC. \nonumber  \\
defeated(D, H, S) & \leftarrow  & s(S), app(D, H, S), \nonumber\\
        & &  neg(H,H_1), \nonumber \\
        & &  h(H_1, S).  \:\:\:  \label{lp:defeat01} \\ 
defeated(D, H, S) & \leftarrow  & s(S), app(D, H, S), \nonumber\\
        & &  neg(H,H_1), \nonumber \\
        & &  believes(H_1, S).  \:\:\:  \label{lp:defeat0} \\ 
defeated(D, H, S) & \leftarrow  & s(S),  app(D, H, S),    \nonumber \\ 
        & &  neg(H,H_1), \nonumber \\ %
        & & app(D_1, H_1, S), \nonumber\\
        & & prefer(D_1, D, S).   \label{lp:defeat}\\ 
ab(D, H, S)& \leftarrow &   s(S), app(D, H, S), \nonumber\\
    & &  neg(H, H_1), app(D_1, H_1, S),  \nonumber \\
    &&   \naf defeated(D_1, H_1, S).  \label{lp:ab}     \\
believes(H, S) & \leftarrow & app(D, H, S), \nonumber \\
        & & \naf ab(D, H, S), \nonumber\\
        & & \naf defeated(D, H, S).  \label{lp:holds}
%    
%defeated(D, H, S)& \leftarrow &  s(S), app(D, H, S), h(\overline{H}, S).\label{lp:defeat0} \\
%    &&  app(D, H, S),  \nonumber \\
%    &&  h(\overline{H}, S). \nonumber \\       
%defeated(D, H, S)& \leftarrow  & s(S),  app(D, H, S), \label{lp:defeat} \\
%    && app(D_1, \overline{H}, S), prefer(D_1, D, S).  \nonumber \\
%    &&  prefer(D_1, D, S).  \nonumber      \\
%h(l_1, S), \ldots, h(l_k, S), \naf ab(d, l, S)   \label{lp:def}\\
%b(l, S) \leftarrow b(l_1, S), \ldots, b(l_k, S), \naf ab(d, l, S)   \label{lp:def_bel}\\
%app(d, l, S) \leftarrow h(l_1, S), \ldots, h(l_k, S) 
%h(H, S)& \leftarrow  app(D, H, S), \naf ab(D, H, S),  \label{lp:holds} \\
% \naf ab(D, H, S), \nonumber \\ 
% & \naf defeated(D, H, S). \nonumber 
\end{eqnarray} 
%} 
%%\fi
%
\iffalse
{
        \[
        \begin{array}{llr}
     app(D, H, S)  \leftarrow &  s(S), default(D, H, M),  \nonumber\\
        & N  = \# count \{L : mbr(L, M)\},  & \label{lp:app} \\
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.8in}h(L, T) \}= N_h,  \\ 
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.44in}believes(L, T) \}= N_b, \\
        & N_h+N_b=N.  \\ 
        ab(D, H, S) \leftarrow  & s(S), app(D, H, S), neg(H,H_1), & (10)\\   %\label{lp:ab}
        &  app(D_1, H_1, S), \\
        & \naf defeated(D_1, H_1, S).        \\
        defeated(D, H, S) \leftarrow  & s(S), app(D, H, S), neg(H,H_1), \\
        & believes(H_1, S).  \:\:\: & (11) \\ %\label{lp:defeat0} 
        defeated(D, H, S) \leftarrow  & s(S),  app(D, H, S), neg(H,H_1), & (12) \\  %\label{lp:defeat}
        & app(D_1, H_1, S), \\
        & prefer(D_1, D, S).   \\
        believes(H, S) \leftarrow & app(D, H, S), \naf ab(D, H, S),\\
        & \naf defeated(D, H, S).  & (13) \\ %\label{lp:holds}
        \end{array}
        \]
        }
%
\fi

@@

 %\setcounter{equation}{13}

 For reasoning about preferences of the form \eqref{pref}, we employ the rules:
% 
{
\begin{equation}\label{lp:pref}
\begin{array}{ll}
prefer(D_1, D_2, S) \leftarrow &  s(S),   prefer(D_1, D_2, M),   \\
     & h\_body(M, S). 
%     & N  =\# count \{L : \\
%     &  \hspace*{.3in} mbr(L, M)\},  \\ 
%     & N_h = \# count \{L : mbr(L, M),  \\
%     &  \hspace*{.3in}  h(L, S) \} , \\
%     & N_n = \# count \{L : mbr(L, M), \\
%     & \hspace*{.3in} believes(L, S) \} , \\
%     & N_h+N_b==N.
      \end{array}
\end{equation} 
}
%
This rule defines when a preference among two defaults can be applied. 
%
%\iffalse 
%Rule~\eqref{lp:app} defines when a default is \emph{applicable}, i.e., when its body is satisfied. The interaction between defaults and rules in the knowledge base are dealt with using rules~\eqref{lp:ab}---\eqref{lp:defeat}. Rule~\eqref{lp:ab} enforces the principle {\bf (SK)}. It says $ab(d, h, s)$ that if there is another default with the conflicting conclusion ($\overline{h}$) 
%which is not defeated then $ab(d, h, s)$ is true. This atom will be used to suppress the conclusion of $d$.   
%\eqref{lp:defeat0} dismisses the applicability of a default if the contrary of its conclusion is already established. On the other hand, rule \eqref{lp:defeat} expresses that a default is defeated if there is a more preferred default with conflicting conclusion that is applicable.  
%
%
% 
%
%\item Rules for reasoning about beliefs  
%\begin{align}
%b(L, S) \leftarrow h(L, S)  \label{lp:belief}   
%\end{align} 
%
%
%\item For interacting between defaults, we have the following rules: 
%
%%\memo{
%%}
%
%\begin{center} 
%  \begin{tikzpicture} 
%%    \tikzstyle{every node}=[draw,shape=circle} 
%%    \node (s1) at (0:0) {$\alert{s_3:}$}   
%%      \draw[fill=blue!20,style=dashed] (-1.0,-1.0) rectangle (3.0,3.0)  ;
%
%      \draw (0,0) circle(0.4) node {$\textcolor{red}{\small s{-}1}$} ;
%      \draw  (2,0) circle(0.4) node  {$\textcolor{red}{\small s}$} ;
%     \draw  (4,0) circle(0.4) node {$\textcolor{red}{\small s{+}1}$} ;  
% 
%     \draw[-,style=dashed] (-1.6,0) -- (-0.4,0 )  ; 
%     \draw[-] (0.4,0) -- (1.6,0)  ;  
%     \draw[-] (2.4,0) -- (3.6,0 )  ; 
%     \draw[-,style=dashed] (4.4,0) -- (5.6,0 )  ; 
% 
%  \end{tikzpicture} 
%\end{center} 
%
%
%\begin{align}\label{lp:ab} 
%ab(D_1, L, S) \leftarrow \: \: &     default(D_1, L, Body_1),   \\
%    &      app(D_1, L, S), \nonumber \\
%    &      default(D_2, \overline{L}, Body_2), \nonumber \\
%    &      app(D_2, \overline{L}, S), \nonumber \\
%    &      \naf app(D_2, S-1).  \nonumber
%\end{align} 
%
%This rule encodes the principle {\bf (P3)}. 
%
%\fi
%
%
%
%% We will assume that things behave in accordance to the law of inertia. So, we will add to the the program the rule 
%%
%% \begin{align}
%% h(l, S+1) \leftarrow h(l, S),  \naf h(\overline{l}, S)   \label{lp:inertial} \\
%% b(l, S+1) \leftarrow b(l, S),  \naf b(\overline{l}, S)   \label{lp:inertial_bel} 
%% \end{align} 
%
%

@@

%
\begin{proposition}
For each consistent $KB$,  there exists no answer set of $\Pi(KB)$ that contains $believes(l, t)$ and $believes(\overline{l},t)$ for some 
literal $l$ and time step $t$. 
\end{proposition} 
%
\noindent
{\bf Proof.}
The proof of this proposition relies on the fact that $believes(l, t)$ is in an answer set $A$ if the logic program  rule \eqref{lp:rule} (or \eqref{lp:holds}) is satisfied by $A$. If a rule \eqref{lp:rule} is satisfied by $A$ then a $KB$ contains a rule $rule(r,l,body)$ and $body$ is satisfied by $A$. Because of the consistency of $KB$, we know that there exists no rule $rule(r', \overline{l}, body')$ such that $body'$ is satisfied by $A$. For, otherwise, $KB$ is inconsistent. This will imply that for every default $default(d, \overline{l}, body')$, if the default is applicable, then the body of the rule \eqref{lp:defeat0} is satisfied and therefore $defeated(d)$ belongs to $A$. Consequently, the body of \eqref{lp:holds} is not satisfied for the default $d$, and hence, $believes(\overline{l}, t)$ cannot belong to $A$.

@@

Intuitively, $KB  \models {+}stm(b,s)@t$ (resp. $KB  \models {-}stm(b,s)@t$) says that at the time step $t$, at the same time step or later than the time step that the statement about $b$ is made, the statement is believed to be true (resp. believed to be false). $KB \not\models {\pm}stm(b,s)@t$ states that there is no information that supports or denies the statement at the time step $t$. Observe that $b$, $t$, and $s$ are constants representing the question whether or not $stm(b,s)$ is believed to be true at step $t$. As such, requiring $believes(b,t)$ to be in every answer set of the program 
does not mean that $believes(b,t)$ belongs to every answer set for every $t$ such that $s \le t \le k$. 
Observe that this entailment can be computed in ASP by the following rules: 
%
{
\begin{eqnarray}
\mathtt{t}(H, T) & \leftarrow & stm(H, S), s(T), T \ge S, \nonumber \\
        &  & believes(H, T). \label{q1} \\ 
\mathtt{f}(H, T) & \leftarrow & stm(H, S), s(T), T \ge S, neg(H,H_1),  \nonumber\\ 
& & believes(H_1, T).  \label{q2}\\
\mathtt{u}(H, T) & \leftarrow & stm(H, S), s(T), T \ge S, neg(H,H_1), \nonumber  \\
& & \naf believes(H, T), \nonumber \\
&&  \naf believes(H_1, T). \label{q3}
\end{eqnarray}
}
%
Let $\Pi_Q = \Pi(KB) \cup \{\eqref{q1}{-}\eqref{q3}\}$. It can be shown that 
 $KB  \models {+}stm(b,s)@t$, $KB  \models {-}stm(b,s)@t$, and  $KB  \models {\pm}stm(b,s)@t$ 
 correspond to $\Pi_Q \models \mathtt{t}(b,t)$, $\Pi_Q \models \mathtt{f}(b,t)$, and 
 $\Pi_Q \models \mathtt{u}(b,t)$, respectively. Hence, we can compute skeptical entailment for the truthfulness of a statement 
using two calls to an ASP-solver.

@@

  \begin{equation}\label{lp:rule:tt}
\begin{array}{l}
h(H, S+T) \leftarrow  s(S),  
        rule(_, H^T, M),  \\ 
     \hspace{.1cm} N = \# count \{L^T : mbr(M, L^T)\},  \\ 
    \hspace{.1cm}  \# count \{L^T : mbr(M, L^T),  h(L, T{+}S) \} {==} N. 
      \end{array}
\end{equation}
